---
title             : "Reproducible Methods for Face Research"
shorttitle        : "Face Methods"

author: 
  - name          : "Lisa DeBruine"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "lisa.debruine@glasgow.ac.uk"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
  - name          : "Iris Holzleitner"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Institute of Neuroscience & Psychology, University of Glasgow"
  - id            : "2"
    institution   : "University of the West of England, Bristol"

authornote: |
  This research was funded by ERC grant #647910 (KINSHIP).

abstract: |
  Face stimuli are commonly created in non-reproducible ways. This paper will introduce the open-access online platform webmorph and its associated R package webmorphR. It will explain the technical processes of morphing and transforming through a case study of creating face stimuli from an open-access image set.
  
keywords          : "faces; morphing; transforming; reproducible; webmorph"
wordcount         : "X"

bibliography      : ["r-references.bib","biblio.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : html_document #papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library("webmorph")
r_refs("r-references.bib")

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```


## Introduction

People manipulate faces. 

Give some examples. 

Scope of this type of research.

### Common Techniques

#### Photoshop/Image editors

- [@Gronenschild_2009]
  
#### Commerical morphing 

* 831 Google Scholar responses for "fantamorph face"
* 158 Google Scholar responses for "WinMorph face"
* Many others: MorphThing (no longer available), xmorph, et. Basically impossible to survey the literature about the methods used because of poor documentaation

#### Codable Methods

* imagemagick
* Matlab
* Psychomorph
* WebMorph
  

### Reproducibility!

I gave up on a research project once because I couldn't figure out how to manipulate spatial scale in MatLab to make my stimuli look like a relevant paper. When I contacted the author, they didn't know how the stimuli were created because a postdoc just did it in photoshop.

Faces are sampled, so replications should sample new faces as well as new participants. 

Difficulty in creating equivalent face stimuli is a barrier to this, resulting in stimulus sets that are used across dozens or hundreds of papers. 

* The Chicago Face Database [@CFD_2015] has been cited in almost 800 papers.
* Ekman POFA **selling** for [$399](https://www.paulekman.com/product/pictures-of-facial-affect-pofa/) for " 110 photographs of facial expressions that have been widely used in cross-cultural studies, and more recently, in neuropsychological research".
* Image sets are often private and reused without clear attribution (FRL and Perception Lab are particularly bad for this).

### Main techniques

#### Averaging



* Visualise group differences

#### Transforming


* Sexual dimorphism




## Methods

### Averaging

### Transforming


## Case Study

### London Face Set

### Delineation

Automatic versus manual delineation.

### Normalisation

Why normalise?

2 point versus Procrustes normalisation (in webmorphR)

```{r procrustes, fig.width = 8, fig.height = 8}
lisa <- faces("lisa")

orig <- plot(lisa, pt.plot = TRUE, labels = "", nrow = 1)

twopt <- align(lisa, pt1 = 63, pt2 = 81, patch = TRUE) %>% 
  plot(pt.plot = TRUE, labels = "", nrow = 1)

# any two points that are standard on the image
# should work for procrustes alignment
lisa_proc <- align(lisa, pt1 = 63, pt2 = 81,
                      procrustes = TRUE, patch = TRUE)

procr <- plot(lisa_proc, pt.plot = TRUE, 
              labels = "", nrow = 1)

cowplot::plot_grid(orig, twopt, procr, nrow = 3, 
                   labels = c("original", "two-point", "procrustes"))
```


### Masking

(effect in masc paper)

### Averaging

Texture/no

### Symmetrising

How this is different from LL/RR mirroring.

### Sexual dimorphism transform

Continuum

### Self-resemblance transform


# Discussion

* head position in 2D images
  * morphometics
  * facefuns
* Natural vs standardised source images
  * right image for the question
* Averaging is N=1


\newpage

# References

We used `r cite_r("r-references.bib")` to produce this manuscript.

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
